---
title: "Descriptive Analysis"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Install and load necessary libraries
#install.packages("vcd")
library(vcd)
#install.packages("corrplot")
library(corrplot)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("GGally")
library(GGally)
#install.packages("gridExtra")
library(gridExtra)
#install.packages("igraph")
library(igraph)

#install.packages("tidyverse")
library(tidyverse)    # all things tidy
#install.packages("haven")
library(haven)

#install.packages("pander")
library(pander)       # nice looking genderal tabulations
#install.packages("furniture")
library(furniture)    # nice Table1() descriptives
#install.packages("texreg")
library(texreg)       # Convert Regression Output to LaTeX or HTML Tables

#install.packages("psych")
library(psych)        # contains some useful functions, like headTail
#install.packages("sjstats")
library(sjstats)      # ICC calculations
#install.packages("sjPlot")
library(sjPlot)       # Visualization for Models

#install.packages("lme4")
library(lme4)         # non-linear mixed-effects models
#install.packages("lmerTest")
library(lmerTest)
#install.packages("effects")
library(effects)      # Effect displays for Models
#install.packages("effectsize")
library(effectsize)
#install.packages("interactions")
library(interactions)
#install.packages("performance")
library(performance)
#install.packages("modeest")
library(modeest)
```

\newpage 

# Descriptive Analysis Plan:

## Objective: Conduct a descriptive analysis of the dataset, focusing on categorical variables and providing summary statistics for the GPA variable.

**Load and Inspect Data:**

* Load the dataset.

* Inspect the structure of the dataset, checking for the data types and missing values.

**Data Cleaning:**

* Ensure that the 'gpa' column is treated as a numeric variable.

**Descriptive Statistics for GPA:**

Compute and present the following summary statistics for the 'gpa' column:

* Mean

* Median

* etc.

**Distribution of GPA:**

* Create a histogram to visualize the distribution of GPA.

**Descriptive Statistics for Categorical Variables:**

* Identify and list all categorical variables in the dataset.

For each categorical variable, compute and present the following summary statistics:

* Number of levels (unique values)

* Frequency distribution of each level

\newpage 

# Data Cleaning

```{r show_col_types = FALSE}
data <- read_csv("data_tidy.csv")
head(data)
```

```{r}
# List of columns to exclude from conversion
columns_to_exclude <- c("id", "timestamp", "field_of_study", "university", "gpa")

# List of columns to convert to categorical
columns_to_convert <- setdiff(names(data), columns_to_exclude)

# Convert selected columns to categorical
data[columns_to_convert] <- lapply(data[columns_to_convert], as.factor)

```
 
\newpage

# Descriptive Statistics for GPA

```{r}
# Select only numeric variables for summary statistics
numeric_data <- data %>%
  select_if(is.numeric)

# Summary statistics for numeric variables
summary(numeric_data)
```
 
\newpage

# Distribution of GPA

```{r, warning=FALSE}
ggplot(data, aes(x = gpa)) +
  geom_histogram(fill = "green", alpha = 0.5, bins = 40) +
  labs(title = "Distribution of GPA", x = "", y = "")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7))
```

\newpage

# Descriptive Statistics for Categorical Variables

```{r}
# Select only categorical variables for frequency table
categorical_data <- data %>%
  select_if(is.factor)

# Frequency table for categorical variables
for (col in colnames(categorical_data)) {
  print(table(categorical_data[[col]], dnn = col))
}

```

\newpage 

# Some Insights

**Preferred Study Environments:**

* Home and Library Dominance: A significant number of respondents prefer studying at home or in the library, suggesting a preference for quiet and controlled environments.

* Campus Spaces: While some students opt for on-campus common spaces, the preference for classrooms, coffee shops, parks, or other outdoor spaces is relatively lower.

**Background Noise Preferences:**

* Diverse Noise Tolerance: Students exhibit diverse preferences for background noise, with a sizable proportion comfortable with some ambient noise. This suggests that designing flexible study spaces catering to various noise preferences could enhance the overall learning environment.

* Varied Study Times: Students have diverse preferences for study times, indicating that institutions should provide flexible scheduling for classes, study sessions, and resources.

**Resource Preferences:**

* Interactive and Multimedia Resources: A notable preference for interactive simulations, videos, and multimedia resources implies that integrating technology and multimedia into teaching methods could be well-received.

* Varied Resource Usage: Preferences for lecture notes, textbooks, and online articles vary, indicating a need for diverse resource availability.

**Consumed Beverages:**

* Caffeine Dominance: Coffee is the most consumed beverage. Consideration of caffeine-friendly spaces and resources may enhance student satisfaction.

**Managing Stress:**

* Diverse Stress Coping Mechanisms: Students employ various stress management methods, including engaging in hobbies, exercising, meditation, and spending time in nature. Institutions could provide resources and support for a variety of stress-relief activities.

**Demographics and Study Patterns:**

* Unemployment and Study Hours: A majority of respondents are unemployed, suggesting that institutions could provide additional support or resources for job placement. Understanding varied study hour patterns can aid in scheduling and resource allocation.

**Attendance and Academic Support:**

* Regular Attendance: The majority attend classes regularly. This signals an engaged student body, but institutions should continue to provide academic support and resources for those who might need additional assistance.

**Academic Goals and Future Plans:**

* Diverse Academic Aspirations: With diverse academic goals and future plans, institutions can tailor support services to cater to students pursuing different paths, whether it be further education, immediate employment, or exploring new interests.




